---
title: 'Physcraper: A Python package for continually updated gene trees'
author:
  - Luna L. Sanchez Reyes
  - Martha Kandziora
  - Emily Jane McTavish
  - University of California, Merced
date: '`r format(Sys.time(), "%d %B, %Y")`'
bibliography: paper.bib
csl: apa.csl
output:
  redoc::redoc
---


# Introduction

Phylogenies are important.

Generating phylogenies is not easy and it is largely artisanal. Although many efforts to automatize the process have been done, and the community is using those more and more, automatization of phylogenetic reconstruction is still not a widespread practice and among other benefits, it might be key for adoption of better reproducibility practices in the phylogenetics community. ***paragraph better to end discussion??? ***

The process of phylogenetic reconstruction implies many steps (that I generalize to the following):

1. Obtention of molecular or morphological character data -- get DNA from some organisms
and sequence it, or get it from an online nucleotide data repository, such as GenBank [@benson2000genbank; @wheeler2000database].
1. Assemble a hypothesis of homology -- Create a matrix of your character data, by
aligning the sequences, in the case of molecular data. Make sure thay are paralogs!
1. Analyse this hypothesis of homology to infer phylogenetic relationships among
the organisms you are studying -- Use different available programs to infer molecular
evolution, trees and times of divergence.
1. Discuss the inferred relationships in the context of previous hypothesis, the
biology and biogeography of the organisms, etc. -- Answer the question, *is this phylogenetic solution fair/reasonable?*

Each of these steps require different types of specialized training: in the field,
in the lab, in front of a computer, discussions with experts in the methods, and/or in the biological group of study.
All of these steps also require considerable amounts of time for training and implementation.

In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
pipelines that mine already available molecular data from the GenBank repository [@benson2000genbank; @wheeler2000database],
to obtain homologous characters that can be used for phylogenetic reconstruction.
These tools have been presented as aid for the nonspecialist to decrease some of
the difficulties in the generation of phylogenetic knowledge. However, they are
not that often used as so, suggesting that there are still difficulties for the nonspecialist.
The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
However, automatizing the assembly of the character data set is a crucial step towards
reproducibility for a task that was otherwise primarily artisanal and hence largely
non-reproducible.

Even if it is hard to obtain phylogenies, we invest copious amounts of time and energy in generating them.
Issues such as food security, global warming, global health are crucial to solve and phylogenies might help.
There is a lot of phylogenetic knowledge already available in published peer-reviewed studies.
In this sense, the non-specialists (and also the specialist) face a new problem: how do I choose the best phylogeny.

Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank [@benson2000genbank; @wheeler2000database].

We present a way to automatize and standardize the comparison of phylogenetic hypotheses and to allow reproducibility of this last step of the research process.

A key aspect of the standard phylogenetic workflow is comparison with already existing
phylogenetic hypotheses and with phylogenies that are considered "best" by experts
not only in phylogenetics, but also experts on the focal group of study.


Concerns I think people have about these tools:
- Errors in identification of sequences
- Little control along the process
- Too much of a black box?

Most of these phylogenies are being constructed by people learning about the methods,
so they want to know what is going on.

The pipelines are so powerful and they will give you an answer, but there
is no way to assess if it is better than previous answers, it just assumes it is
better because it used more data.

All these pipelines start tree construction from zero? Yes.

The goal of Physcraper is to build upon previous phylogenetic knowledge,
allowing a direct comparison between existing phylogenies and phylogenies that are constructed
using new genetic data retrieved from a public nucleotide database (i.e., GenBank [@benson2000genbank; @wheeler2000database]).

To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
to the TreeBase database, to (1) get the original DNA data set matrices (alignments) that produced
a phylogeny that was published and then made available in the OToL database, (2)
use this DNA alignments as a starting point to get new genetic data belonging
to the focal group of study, to (3) finally update the phylogenetic relationships in the group.

A less automated workflow is one in which the alignments that generated the published
phylogeny are stored in other public database (such as DRYAD) or elsewhere (the users computer), and
are provided by the users.

The original tree is by default used as starting tree for the phylogenetic searches, but it
can also be set as a full topological constraint or not used at all, depending on
the goals of the user.

Physcraper implements node by node comparison of the the original and the updated trees, using the conflict API of OToL.

# How does Physcraper work?

## The input: a study tree and an alignment

- The study tree is a published phylogenetic tree stored in the OToL database, phylesystem [@mctavish2015phylesystem]. The main
  reason for this is that trees in phylesystem have a set of user defined characteristics
  that are essential for automatizing the phylogeny update process. The most relevant of these being the definition of ingroup and outgroup. Outgroup and ingroup taxa in the original tree are identified and tagged. This allows to automatically set the root for the updated tree on the next steps of the pipeline.
  A user can choose from the 'r rotl::tol_about()$num_source_trees' published trees supporting the resolved node of the synthetic tree in the OToL website (<>). If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (<https://tree.opentreeoflife.org/curator).
- The alignment should be a gene alignment that was used to generate the tree. The original
  alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
  DRYAD (<http://datadryad.org/>), or the journal were the tree was originally published.
  If the alignment is stored in TreeBase, `physcraper` can download it directly,
  either from the TreeBASE website (<https://treebase.org/>)
  or through the TreeBASE GitHub repository (SuperTreeBASE; <https://github.com/TreeBASE/supertreebase>).
  If the alignment is on another repository, or provided personally by the owner, a copy of it has to be
  downloaded by the user, and it's local path has to be provided as an argument.
- A taxon name matching step is performed to verify that all taxon names on the tips
  of the tree are in the DNA character matrix and vice versa.
- A ".csv" file with the summary of taxon name matching is produced for the user.
- Unmatched taxon names are dropped from both the tree and alignment.
  <!--The minimum amount of tips necessary for a physcraper run is just one!!!-->
  Technically, just one matching name is needed to perform the searches. Please, see next section.
- A ".tre" file and a ".fas" file containing only the matched taxa are generated and saved in the `inputs` folder to be used in the following steps.

## DNA sequence search and cleaning

- The next step is to identify the search taxon within the reference taxonomy. The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group. Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy. The search taxon can be provided by the user. If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
  matched taxa belonging to the ingroup in the tree, that is also a named clade in the reference taxonomy. This is known as the Most Recent Common
  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA).
  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
  To automatically identify the MRCAT of a group of taxon names, we make use of the OToL taxonomy tool (<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>).

  Users can provide a search taxon that is either a more or a less inclusive
  clade relative to the ingroup of the original phylogeny. If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
  the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
- The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; altschul1997gapped] is used to identify
  similarity between DNA sequences within the search taxon in a nucleotide
  database, and the accepted sequences on the alignment.
  The blastn function from the BLAST command line tools [@camacho2009blast] is used for local-database searches.
  A modified biopython blast function is used for web-based searches.
  <!--***How does the blast is setup? Is it an all-blast-all??***-->
- The DNA sequence similarity search can be done on a local database that is easily
  setup by the user. In this case, the blastn function is used to performs the similarity search [@camacho2009blast].
- The search can also be performed remotely, on the NCBI database. In this case, the
  bioPython BLAST function was modified to accepts is used to perform the similarity search.
- A pairwise alignment-against-all BLAST search is performed. This means that each sequence
  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
  along with their corresponding identification numbers (accesion numbers in the case of the GenBank database). This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
- Matched sequences below an e-value, percentage similarity, and outside a minimum
  and maximum length threshold are discarded. ***REPORT THE DEFAULT VALUES AND DESCRIBE WHAT THEY MEAN*** This filtering leaves out genomic sequences.
  All acepted sequences are asigned an internal identifier, and are further filtered.
- Because the original alignments usually lack database id numbers, a filtering
  step is needed. Accepted sequences that belong to the
  same taxon of the query sequence, and that are either identical or shorter than
  the original sequence are discarded. Only longer sequences belonging to the
  same taxon as the orignal sequence will be considered further for analysis.
- Among the remaining filtered sequences, there are usually several exemplars per taxon.
  Although it can be useful to keep some of them to, for example, investigate monophyly
  within species, there can be hundreds of exemplar sequences per taxon for some markers.
  To control the number of sequences per taxon in downstream analyses,
  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
- Reverse complement sequences are identified and translated.
- Users can choose to perform a more "cycles" of sequence similarity search, by blasting the newly found sequences. This can be done iteratively, but by default only sequences in the alignment are blasted. ***Is there an argument to control the number of cycles of blast searches with new sequences?***
- Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (physcraper/taxonomy), so they are accesible for further runs.
- A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.

## DNA sequence alignment

- The software MUSCLE [@edgar2004muscle] is implemented to perform alignments.
- First, all new sequences are aligned using default MUSCLE options.
- Then, a MUSCLE profile alignment is performed, in which the original alignment
  is used as a template to align new sequences. This ensures that the final alignment
  follows the homology criteria established by the original alignment.
- The final alignment is not further processed automatically. We encourage users
  to check it either by eye and perform manual refinement or using any of the many
  tools for alignment processing, to eliminate columns with no information.

## Tree reconstruction and comparison

- A gene tree is reconstructed for each alignment provided, using a Maximum Likelihood approach implemented with the software RAxML [@stamatakis2014raxml]
  with 100 classic rapid bootstrap [@felsenstein1985confidence] replicates by default. The number of bootsrap replicates can be modified by the user.
  Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in @lemoine2018renewing.
- Bootstrap results are summarized with Dendropy ADD CITATION
- The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
- Tips on all trees generated by physcraper are defined by a taxon name space, allowing to perform comparisons and conflict analyses.
- Robinson Foulds weighted and unweighted metrics ARE CALCULATED WITH DENDROPY TOO.
- Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to CITE REDELINGS AND HOLDER [@redelings and holder]
- For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
- A SUGGESTED DEFAULT ROOTING BASED ON THE OPEN TREE TAXONOMY is implemented for now.  DESCRIBE HOW IT WORKS. SAY THAT IT IS A PROBLEM. Automatic rooting is not that smart yet. The best way right now is for users to define outgroups so trees are better rooted.
- Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon. This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
- Conflict information can only be generated in the context of the whole Open Tree of
  Life. Otherwise, it is not really possible to get conflict data.
***- One way to compare two independent phylogenetic trees is to compare them both to
the synthetic OToL and then measure how well they do against each other***

# Examples

## The hollies

The genus *Ilex* is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
It encompasses between 400-600 living species. A review of litterature shows that there are three published phylogenetic trees, showing relationships within the hollies.
The first one has been made available both on OToL phylesystem and synth tree, and on treeBASE, it samples 48 species.
The second has not been made available anywhere, not even in supplementary data of the journal.
***Contact authors? They seem old school, probably do not wanna share their data.***
The most recent one has been made available in the OToL Phylesystem and DRYAD. It is the best sampled yet, with 200 species. However,
it has not been added to the syntehtic tree yet.
This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database [@benson2000genbank; @wheeler2000database]. Updating the oldest tree, we should get something very similar to the newest tree.


## The Ascomycota


Let's be more specific now about our X group and say it is the Ascomycota.
The best tree currently available in OToL was published by @schoch2009ascomycota.
The first step, is to get the Open Tree of Life study id. There are some options to do this:
- You can go to the Open Tree of Life website and browse until you find it, or
- you can get the study id using R tools:
  - By using the TreeBase ID of the study (which is not fully exposed on the
    TreeBase website home page of the study, so you have to really look it up manually):
```{r, treebase, eval = TRUE}
rotl::studies_find_studies(property = "treebaseId", value = "S2137")
```
  - By using the name of the focal clade of study (but this behaved very differently):
```{r, focal, eval = FALSE}
rotl::studies_find_studies(property="ot:focalCladeOTTTaxonName", value="Ascomycota")
```

Once we have the study id, we can gather the trees published on that study:
```{r, treeids, eval = TRUE}
rotl::get_tree_ids(rotl::get_study_meta("pg_238"))
rotl::candidate_for_synth(rotl::get_study_meta("pg_238"))
my_trees <- rotl::get_study("pg_238")
```
Both trees from this study have `r tryCatch(ape::Ntip(my_trees[[1]]), error = function(e){return(NA)})` tips.

Let's check what one of the trees looks like:

<!-- ```{r fungi-phylo, eval = TRUE, fig.retina = 2, fig.width = 2} -->
<!-- ape::plot.phylo(ape::ladderize(my_trees[[1]]), type = "phylogram", cex = 0.1) -->
<!-- ``` -->

1. Download the alignment from TreeBase
If you are on the TreeBase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to TreeBase and to the Open Tree of Life repository.
To make this task easier, you can use a command to download everything into your working folder:
```
physcraper_run.py -s pg_238 -t tree109 -o ../physcraper_example/pg_238
```
In this example, all alignments posted on TreeBase were used to reconstruct both trees.

1. With the study id and the alignment files saved locally, we can do a physcraper run with the command:

```
physcraper_run.py -s pg_238 -t tree109 -a treebase_alns/pg_238tree109.aln -as "nexus" -o pg_238
```

## Testudines example

Phylogeny of the Testudines 6 tips from @crawford2012more
There is just one tree in OToL.
There is just one alignment on [treebase](https://treebase.org/treebase-web/search/study/matrices.html?id=12742) with all the 1 145 loci.


```
physcraper_run.py -s pg_2573 -t tree5959 -tb -db ~/branchinecta/local_blast_db/ -o pg_2573
```
# Discussion

Data repositories hold more information than meets the eye.
Besides the actual data, they have other types of information that can be used for the advantage of science.

Usually, initial ideas about the data are changed by analyses.
We expect that this new ideas on the data can be registered on data bases,
exposing new comers to expert understanding about the data.

There are many tools that are making use of DNA data repositories in different ways.
Most of them focus on efficient ways to mine the data -- getting the most homologs.
Some focus on accurate ways of mining the data - getting real and clean homologs.
Others focus on refinement of the alignment.
Most focus on generating full trees *de novo*, mainly for regions of the Tree of
Life that have no phylogenetic assessment yet in published studies, but also for
regions that have been already studied and that have phylogenetic data already.

All these tools are great efforts for advancing towards reproducibility in phylogenetics,
a field that has been largely recognised as somewhat artisanal.
We propose adding focus to other sources of information available from data repositories.
Taking advantage of public DNA data bases have been the main focus. However, phylogenetic knowledge is
also accumulating fast in public and open repositories.
In this way, the physcraper pipeline can be complemented with other tools that have
been developed for other purposes.

We emphasize that physcraper takes advantage of the knowledge and intuition of the expert
community to build upon phylogenetic knowledge, using not only data accumulated in
DNA repositories, but phylogenetic knowledge accumulated in tree repositories.
This might help generate new phylogenetic data. But physcraper does not seek to generate full phylogenies *de novo*.

Describe again statistics to compare phylogenies provided by physcraper via OpenTreeOfLife.
Mention statistics provided by other tools: PhyloExplorer [@ranwez2009phyloexplorer].
Compare and discuss.

How is physcraper already useful:
- to mine targeted sequences, in this way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota does not do baited analyses, I think, only clustered analyses.
- Finding

How can it be used for the advantage of the field:
- rapid phylogenetic placing of newly discovered species, as mentioned in @webb2010biodiversity
- obtain trees for ecophylogenetic studies, as mentioned in @helmus2012phylogenetic
- one day could be used to sistematize nucleotide databases, such as Genbank [@benson2000genbank; @wheeler2000database], as mentioned in @san2010molecular, i.e., curate ncbi taxonomic assignations.
- allows to generate custom species trees for downstream analyses, as mentioned in @stoltzfus2013phylotastic


Things that physcraper does not do:
- analyse the whole GenBank database [@benson2000genbank; @wheeler2000database] to find homolog regions suitable to reconstruct phylogenies, as mentioned in @antonelli2017toward. There are already some very good tools that do that.
- provide basic statistics on data availability to assemble molecular datasets, as mentioned by @ranwez2009phyloexplorer. Phyloexplorer does this?
- it is not a tree repo, as phylota is, mentioned in @deepak2014evominer

## Tools that automatize any part of the process of phylogenetic reconstruction:

### 1. Mining DNA databases to generate datasets suitable for phylogenetic reconstruction

| Tool | Citation | Cited by | Description | Supermatrix/gene tree/species tree |
|-----|------|:------:|:------:|:------:|
| Phylota | @sanderson2008phylota | 122 studies | finds sets of DNA homologs on the GenBank database; phylogenetic reconstruction | Supermatrix |
| AMPHORA | @wu2008simple | 458 studies | baited search; protein markers on phylogenomic data; personal database of genomes or metagenomic data, manually downloaded either from a public database or from private data; phylogenetic reconstruction | Supermatrix |
| PHLAWD | @smith2009mega | 234 studies | Baited search of DNA markers on the GenBank database; phylogenetic reconstruction | Supermatrix |
| Unnamed [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals), only available from [supplementary data](https://static-content.springer.com/esm/art%3A10.1186%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP) of the journal | @peters2011taming | 64 studies | mining public DNA databases, focuses on filtering massive amounts of mined sequences by using established "criteria of compositional homogeneity and defined levels of density and overlap" | Supermatrix |
| Unnamed | @grant2014building | 38 studies | predecessor of phylotol; homolog clustering; public and/or personal DNA database; phylogenetic reconstruction; broad taxon analyses; remove contaminant sequences, based on similarity and on phylogenetic position | supermatrix |
| Unnamed | @chesters2014protocol | 10 studies | algorithm that mines GenBank data to delineate species in the insecta. The authors present a nice comparison with the phylota algorithm | Species trees?? |
| PUmPER | @izquierdo2014pumper | 14 studies | perpetual updating with newly added sequences to GenBank | not sure yet |
| DarwinTree | @meng2015darwintree | 6 studies | predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP), takes data from GenBank, EMBL and DDBJ for land plants only| not sure |
| NCBIminer | @xu2015ncbiminer | 4 studies | part of darwintree | not sure |
|SUMAC | @freyman2015sumac | 19 studies | both “baited” analyses and single‐linkage clustering methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix | not sure |
|STBase | @mcmahon2015stbase | 7 studies | pipeline for species tree construction and the public database of one million precomputed species trees | species trees |
| Unnamed | @papadopoulou2015automated | 17 studies | Automated DNA-based plant identification for large-scale biodiversity assessment | not sure |
| BIR | @kumar2015bir | 6 studies | blast, align, identify homologs via constructed trees, curate and realign | supermatrix |
| SUPERSMART | @antonelli2017toward | 35 studies | baited analyses up to bayesian divergence time estimation | supermatrix |
| SOPHI | [@chesters2017construction | 17 studies | Searches DNA sequence data from repos other than GenBank, such as transcriptomic and barcoding repos | not sure |
| phyloSkeleton | @guy2017phyloskeleton | 5 studies | focuses on taxon sampling; baited genomic sequences; public database (NCBI and JGI); marker identification| supermatrix |
| OneTwoTree | @drori2018onetwotree | 7 studies | Web‐based, user-friendly, online tool for species-tree reconstruction, based on the *supermatrix paradigm* and retrieves all available sequence data from NCBI GenBank| supermatrix |
| pyPhlawd | @smith2019pyphlawd | 6 studies | baited and clustering analyses | Supermatrix or gene tree |
| Phylotol | @ceron2019phylotol | 5 studies | "phylogenomic pipeline to allow easy incorporation of data from  high-throughput sequencing studies, to automate production of both multiple sequence alignments and gene trees, and to identify and remove contaminants. PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree of life", i.e., bacteria and unicellular eukaryotes | supermatrix and gene trees |
| phylotaR | @bennett2018phylotar | studies |  |  |

According to @ceron2019phylotol, PhyLoTA and BIR "focus on the identification and collection
of homologous and paralog genes from public databases such as GenBank", while both AMPHORA and PHLAWD
"focus on the construction and refinement of robust alignments rather than the collection of homologs."

### 2. Searching phylogenetic tree databases

PhyloFinder [@chen2008phylofinder] - cited by 18: a search engine for phylogenetic databases, using
trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies

### 3. Mining phylogenetic tree databases

PhyloExplorer [@ranwez2009phyloexplorer] - cited by 21: a python and MySQL based website to facilitate
assessment and management of phylogenetic tree collections. It provides "statistics describing the collection,
correcting invalid taxon names, extracting taxonomically relevant parts of the collection
using a dedicated query language, and identifying related trees in the TreeBASE database".

### 4. Pipeline for phylogenetic reconstruction

PhySpeTre [@fang2019physpetree] - no citations yet - no sequence retrieval, just phylogenetic reconstruction
pipeline.

### 5. getting metadata and not sequences from GenBank.

Datataxa @ruiz2019datataxa - no citations yet - focus on extracting metadata from GenBank sequence information.


## Phylota overview

```{r phylota, child="100-phylota-overview.Rmd"}
```

# Acknowledgements

We acknowledge contributions from

The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No. ACI-1429783).

# Authors' Contributions

# Data Avilability

# References
