---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'Physcraper: A Python package for continually updated gene trees'
tags:
  - Python
  - phylogenetics
  - update
  - species tree
  - Open Tree
authors:
  - name: Luna L. Sanchez Reyes
    orcid: 0000-0000-0000-0000
    affiliation: 1 # (Multiple affiliations must be quoted, like this: "1, 2")
  - name: Emily Jane McTavish
    orcid: 0000-0000-0000-0000
    affiliation: 1
affiliations:
 - name: University of California, Merced
   index: 1
 - name: Institution 2
   index: 2
citation_author: Sanchez-Reyes et. al.
date: 24 April 2020
year: 2020
bibliography: paper.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---

# Summary

Update trees that were reported on a published peer-reviewed study, uploaded to Treebase and to the Open Tree of Life database

# Examples

Updating trees from @schoch2009ascomycota

1. Get study id from Open Tree of Life using R
  - Getting the treebase ID (which is not fully exposed on the Treebase website home page of the study, so you have to look it up manually)
```{r}
rotl::studies_find_studies(property = "treebaseId", value = "S2137")
```

1. Download the alignment from Treebase
  - If you are on the Treebase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to Treebase and to the Open Tree of Life repository.

```
```

```{bash}
opentree_scrape.py -s pg_238 -t tree109 -a treebase_alns/pg_238-LSU-M3956.nex -as "nexus" -o tmp
```


# Citations

Tools that do similar things:
pyPhlawd [@smith2019pyphlawd]


For a quick reference, the following citation commands can be used:
- `@author:2001`  ->  "Author et al. (2001)"
- `[@author:2001]` -> "(Author et al., 2001)"
- `[@author1:2001; @author2:2001]` -> "(Author1 et al., 2001; Author2 et al., 2002)"

# Rendered R Figures

Figures can be plotted like so:

```{r}
plot(1:10)
```

# Acknowledgements

We acknowledge contributions from Brigitta Sipocz, Syrtis Major, and Semyeong
Oh, and support from Kathryn Johnston during the genesis of this project.

# References
