---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'Physcraper: A Python package for continually updated gene trees'
tags:
  - Python
  - phylogenetics
  - update
  - species tree
  - Open Tree
authors:
  - name: Luna L. Sanchez Reyes
    orcid: 0000-0000-0000-0000
    affiliation: 1 # (Multiple affiliations must be quoted, like this: "1, 2")
  - name: Emily Jane McTavish
    orcid: 0000-0000-0000-0000
    affiliation: 1
affiliations:
 - name: University of California, Merced
   index: 1
 - name: Institution 2
   index: 2
citation_author: Sanchez-Reyes et. al.
date: 24 April 2020
year: 2020
bibliography: paper.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  eval = TRUE
)
```

# Summary

1. Phylogenies are still a key part of research in all areas of biology. The nonspecialist
can use one of the many tools available to automatize phylogenetic reconstruction.
However, interpretation of results, comparison with previously available phylogenetic
hypotheses, and choosing of one phylogeny for discussion still impose difficulties
to one that is not a specialist on phylogenetic methods or on a particular group of study.

1. Physcraper is an open‐source Python program that automatizes the update of public
phylogenies by making use of public DNA sequence data and taxonomic information.

1. Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
hypothesis based on public expert phylogentic knowledge.
The phylogenetics and group specialist will find it useful as a tool to facilitate comparison
of alternative phylogenetic hypotheses (topologies).
***Is physcraper intended for the nonspecialist?? We have two types of nonspecialists:
the ones that do not know about phylogenetic methods and the ones that might know
about phylogenetic methods but do not know of a certain biological group.***

1. Physcraper implements node by node comparison of the the original and the updated
trees using the conflict API of OToL.

1. We hope the physcraper workflow demonstrates the benefits of open results in phylogenetics and encourages researchers to strive for better data sharing practices.

1. Physcraper can be used with any OS. Detailed instructions for installation and
use are available at <https://github.com/McTavishLab/physcraper>.


# Intro

There is a lot of phylogenetic knowledge already available.
Some of it has been made public at OToL.
Most or all public phylogenies have the potential to be updated with the ever increasing
amount of genetic data that is available on GenBank.
There are now various methods to automatize use of GenBank data for phylogenetic reconstruction, which are
generally intended to facilitate the use of phylogenetic knowledge for the nonspecialist.
However, none of them provide a means to compare with previously available expert
knowledge on a particular group.

A key aspect of the standard phylogenetic workflow is comparison with already existing
phylogenetic hypotheses and with phylogenies that are considered "best" by experts
not only in phylogenetics, but also experts of the focal group of study.

It is well known that GenBank holds enormous amounts of genetic data, and it continues to grow.
A lot of this genetic data has the potential to be used to reconstruct the phylogenetic
history of various organisms [@sanderson2008phylota].
Pipelines that harness this potential have been available for over a decade now, such
as the Phylota brower, and PHLAWD. New ones keep on being developed, such as SUPERSMART
and the upgraded version of PHLAWD, PyPHLAWD.
Notably large phylogenies have been constructed using some of these tools,
Some other have not been used that much. So, how well accepted is this approach in the community?

Concerns with these tools:
Errors in identification of sequences
Little control along the process
Too much of a black box?

Most of these phylogenies are being constructed by people learning about the methods,
so they want to know what is going on.

The pipelines are so powerful and they will give you an answer, but there
is no way to assess if it is better than previous answers, it just assumes it is
better because it used more data.

All these pipelines start tree consruction from zero?

The goal of Physcraper is to build upon previous phylogenetic knowledge,
allowing a direct comparison of existing phylogenies to phylogenies that are constructed
using new genetic data from GenBank

To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
to the TreeBase database, to (1) get the original DNA dataset matrices (alignments) that produced
a phylogeny that was published and then made available in the OToL database, (2)
use this DNA alignments as a starting point to get new genetic data belonging
to the focal group of study, to (3) finally update the phylogenetic relationships in the group.

A less automated workflow is one in which the alignments that generated the published
phylogeny are stored in other public database (such as DRYAD) or elsewhere (the users computer), and
are provided by the users.

The original tree is by defualt used as starting tree for the phylogenetic searches, but it
can also be set as a full topological constraint or not used at all, depending on
the goals of the user.

Physcraper implements node by node comparison of the the original and the updated trees, using the conflict API of OToL.

# Physcraper overview



# Use case/ example

Imagine you are starting to work on a new biological group X. You have not much of an idea about
its phylogenetic relationships, you are a newly established researcher, and the group
is not anything any of your collaborators have worked on before.
A good idea is to start an intensive litt review on the phylogenetics of the group.
Rapidly, you find out there are 5 different phylogeneis, that used different markers, and that
the papers, published at different times, do not discuss which phylogeny is
the one accepted by the expert community on X. You might need
to go to the annual conference of X, and even then, you might only find
different and contrasting opinions.
Somewhere along these months or even years doing this task, you looked into the
the OToL database. You found in there some or all the published trees of X, along
with a tree that has been deemed the best tree by curators and ideally experts on X?

## Ascomycota Example

Let's be more specific now about our X group and say it is the Ascomycota.
The best tree currently available in OToL was published by @schoch2009ascomycota.
The first step, is to get the Open Tree of Life study id. There are some options to do this:
- You can go to the Open Tree of Life website and browse until you find it, or
- you can get the study id using R tools:
  - By using the treebase ID of the study (which is not fully exposed on the
    Treebase website home page of the study, so you have to really look it up manually):
```{r, treebase, eval = TRUE}
rotl::studies_find_studies(property = "treebaseId", value = "S2137")
```
  - By using the name of the focal clade of study (but this behaved very differently):
```{r, focal, eval = FALSE}
rotl::studies_find_studies(property="ot:focalCladeOTTTaxonName", value="Ascomycota")
```

Once we have the study id, we can gather the trees published on that study:
```{r, treeids, eval = TRUE}
rotl::get_tree_ids(rotl::get_study_meta("pg_238"))
rotl::candidate_for_synth(rotl::get_study_meta("pg_238"))
my_trees <- rotl::get_study("pg_238")
```
Both trees from this study have `r ape::Ntip(my_trees[[1]])` tips.

Let's check what one of the trees looks like:

```{r fungi-phylo, eval = TRUE, fig.retina = 2, fig.height = 4}
ape::plot.phylo(my_trees[[1]], type = "fan", cex = 0.1)
```

1. Download the alignment from Treebase
If you are on the Treebase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to Treebase and to the Open Tree of Life repository.
To make this task easier, you can use a command to download everything into your working folder:
```
physcraper_run.py -s pg_238 -t tree109 -o ../physcraper_example/pg_238
```
In this example, all alignments posted on Treebase were used to reconstruct both trees.

1. With the study id and the alignment files saved locally, we can do a physcraper run with the command:

```
physcraper_run.py -s pg_238 -t tree109 -a treebase_alns/pg_238tree109.aln -as "nexus" -o pg_238
```

## Testudines example

Phylogeny of the Testudines 6 tips from @crawford2012more
There is just one tree in OToL.
There is just one alignment on [treebase](https://treebase.org/treebase-web/search/study/matrices.html?id=12742) with all the 1 145 loci.

```
physcraper_run.py -s pg_2573 -t tree5959 -tb -db ~/branchinecta/local_blast_db/ -o pg_2573
```


# Tools on a similar track:

Tools that do similar things:
pyPhlawd [@smith2019pyphlawd]
SUPERSMART [@antonelli2017toward]



# Acknowledgements

We acknowledge contributions from

# References
